Follow-up Samples

### 2017-02-09 submission/manuscript_outline_20161024.Rmd
better flow for results

***Common OTUs to both Models***
The reduced models were built based on the most important variables to either classification model.  For the lesion model there were a total of `r length(lesion_top10_varsMDA$variable)` variables [Figure S2] whereas for the before sample model there were a total of `r length(IF_top10_varsMDA$variable)` variables [Figure S3].  For both models FIT resulted in the largest decrease in MDA [Figure S2a & S3a].  There were a total of `r length(common_vars_summary$otu)` OTUs that were common to both models.  Of these OTUs the most common taxonomic identifications were to  `r common_vars_summary['Otu000012', 'lowest_ID']`, `r common_vars_summary['Otu000005', 'lowest_ID']`, `r common_vars_summary['Otu000082', 'lowest_ID']`, and `r paste(gsub("_unclassified", "", common_vars_summary['Otu000006', 'lowest_ID']))`.  The majority of these OTUs had classifications to bacteria typically thought of as commensal [Table S2].

***Positive Probability after Lesion Removal***
Regardless of model used there was a significant decrease in the positive probability of either the sample being lesion or a before sample on follow up [Figure 4 & S4] (full lesion P-value = `r format(filter(all_probs_paired_wilcox, model_type =="lesion" & comparison == "lesion")[, "BH_correction"], scientific = TRUE, digits = 3)`, reduced lesion P-value = `r format(filter(all_probs_paired_wilcox, model_type =="red_lesion" & comparison == "lesion")[, "BH_correction"], scientific = TRUE, digits = 3)`, initial and follow up P-value = `r format(filter(all_probs_paired_wilcox, model_type =="IF" & comparison == "lesion")[, "BH_correction"], scientific = TRUE, digits = 3)`, and reduced initial and follow up P-value = `r format(filter(all_probs_paired_wilcox, model_type =="red_IF" & comparison == "lesion")[, "BH_correction"], scientific = TRUE, digits = 3)`).

For the full and reduced lesion model there was a significant difference in the classification for the lesion model between predicted and actual (P-value = `r format(filter(all_pred_stat_summary, samples_tested == "all" & model_type == "lesion" & info == "McnemarPValue")[, "all"], scientific = TRUE, digits = 3)` and `r format(filter(all_pred_stat_summary, samples_tested == "all" & model_type == "red_lesion" & info == "McnemarPValue")[, "all"], scientific = TRUE, digits = 3)`, respectively) but not for the before sample model (P-value = `r format(filter(all_pred_stat_summary, samples_tested == "all" & model_type == "IF" & info == "McnemarPValue")[, "all"], digits = 3, nsmall = 2)` and `r format(filter(all_pred_stat_summary, samples_tested == "all" & model_type == "red_IF" & info == "McnemarPValue")[, "all"], digits = 3, nsmall = 2)`).  However, the lesion model correctly kept the one individual who still had a carcinoma on follow up above the cut off threshold [Figure 4a & S4a] for a positive call while the before sample models did not [Figure 4b & S4b].

***Treatment and Time Differences***
There was no difference in the amount of change in positive probability for either the full or reduced lesion model for either chemotherapy (P-value = `r format(chemo_rad_summary['chemo_les', 'bh'], digits = 3)` and `r format(chemo_rad_summary['chemo_red_les', 'bh'], digits = 3)`) or radiation therapy (P-value = `r format(chemo_rad_summary['rad_les', 'bh'], digits = 3)` and `r format(chemo_rad_summary['rad_red_les', 'bh'], digits = 3)`). Although the before sample model was similar there was a significant decrease in positive probability for those treated with chemotherapy (P-value = `r format(chemo_rad_summary['chemo_IF', 'bh'], scientific = TRUE, digits = 3)` and `r format(chemo_rad_summary['chemo_red_IF', 'bh'], scientific = TRUE, digits = 3)`).  There was a significant difference in time elapsed in the collection of the follow up sample between adenoma and carcinoma samples (uncorrected P-value = `r format(time_pvalues["time", "uncorrected_p_value"], scientific = TRUE, digits = 3)`) with time passed being lower for adenoma (`r format(time_summary['adenoma', 'mean_days'], digits = 3)` +/- `r format(time_summary['adenoma', 'sd_days'], digits = 3)` days) then carcinoma (`r format(time_summary['carcinoma', 'mean_days'], digits = 3)` +/- `r format(time_summary['carcinoma', 'sd_days'], digits = 3)` days).

>>>>

***Common OTUs to both Models***
The reduced models were built based on the most important variables to either classification model.  For the lesion model there were a total of `r length(lesion_top10_varsMDA$variable)` variables [Figure S2] whereas for the initial sample model there were a total of `r length(IF_top10_varsMDA$variable)` variables [Figure S3].  For both models FIT resulted in the largest decrease in MDA [Figure S2a & S3a].  There were a total of `r length(common_vars_summary$otu)` OTUs that were common to both models.  Of these OTUs the most common taxonomic identifications were to  `r common_vars_summary['Otu000012', 'lowest_ID']`, `r common_vars_summary['Otu000005', 'lowest_ID']`, `r common_vars_summary['Otu000082', 'lowest_ID']`, and `r paste(gsub("_unclassified", "", common_vars_summary['Otu000006', 'lowest_ID']))`.  The majority of these OTUs had classifications to bacteria typically thought of as commensal [Table S2].

***Positive Probability after Lesion Removal***
If there were common OTUs for individuals with adenoma and carcinoma that were different versus normal controls we would expect to find a decrease in the positive probability of the follow up sample to be either lesion or a before sample. This is what we observed regardless of model used (lesion or before sample) or whether it was built on the full or reduced variable data set [Figure 4 & S4] (full lesion P-value = `r format(filter(all_probs_paired_wilcox, model_type =="lesion" & comparison == "lesion")[, "BH_correction"], scientific = TRUE, digits = 3)`, reduced lesion P-value = `r format(filter(all_probs_paired_wilcox, model_type =="red_lesion" & comparison == "lesion")[, "BH_correction"], scientific = TRUE, digits = 3)`, initial and follow up P-value = `r format(filter(all_probs_paired_wilcox, model_type =="IF" & comparison == "lesion")[, "BH_correction"], scientific = TRUE, digits = 3)`, and reduced initial and follow up P-value = `r format(filter(all_probs_paired_wilcox, model_type =="red_IF" & comparison == "lesion")[, "BH_correction"], scientific = TRUE, digits = 3)`).

Although there was a decrease in positive probability for all models there were differences on whether the models classified the samples as still having lesion.  This was most evident for the full and reduced lesion model where there was a large difference between predicted and actual (P-value = `r format(filter(all_pred_stat_summary, samples_tested == "all" & model_type == "lesion" & info == "McnemarPValue")[, "all"], scientific = TRUE, digits = 3)` and `r format(filter(all_pred_stat_summary, samples_tested == "all" & model_type == "red_lesion" & info == "McnemarPValue")[, "all"], scientific = TRUE, digits = 3)`, respectively).  However, the before sample model did not have this disparity between predicted and actual (P-value = `r format(filter(all_pred_stat_summary, samples_tested == "all" & model_type == "IF" & info == "McnemarPValue")[, "all"], digits = 3, nsmall = 2)` and `r format(filter(all_pred_stat_summary, samples_tested == "all" & model_type == "red_IF" & info == "McnemarPValue")[, "all"], digits = 3, nsmall = 2)`). Even though the lesion model varied greatly between predicted and actual it was able to correctly keep the one individual who still had a carcinoma on follow up above the cut off threshold [Figure 4a & S4a] for a positive call while the before sample models did not [Figure 4b & S4b].

***Treatment and Time Differences***
After observing these changes in positive probability We wanted to assess whether additional treatments, such as chemotherapy and radiation, could have an impact on the results that we observed.  There was no difference in the amount of change in positive probability for either the full or reduced lesion model for either chemotherapy (P-value = `r format(chemo_rad_summary['chemo_les', 'bh'], digits = 3)` and `r format(chemo_rad_summary['chemo_red_les', 'bh'], digits = 3)`) or radiation therapy (P-value = `r format(chemo_rad_summary['rad_les', 'bh'], digits = 3)` and `r format(chemo_rad_summary['rad_red_les', 'bh'], digits = 3)`). Although the before sample model was similar there was a significant difference in decreased positive probability for those treated with chemotherapy (P-value = `r format(chemo_rad_summary['chemo_IF', 'bh'], scientific = TRUE, digits = 3)` and `r format(chemo_rad_summary['chemo_red_IF', 'bh'], scientific = TRUE, digits = 3)`).  Finally, we wanted to know if the length of time between the initial and follow up sample could be a possible confounder.  Within our study there was a significant difference in the time elapsed in the collection of the follow up sample between adenoma and carcinoma samples (uncorrected P-value = `r format(time_pvalues["time", "uncorrected_p_value"], scientific = TRUE, digits = 3)`) with time passed being lower for adenoma (`r format(time_summary['adenoma', 'mean_days'], digits = 3)` +/- `r format(time_summary['adenoma', 'sd_days'], digits = 3)` days) then carcinoma (`r format(time_summary['carcinoma', 'mean_days'], digits = 3)` +/- `r format(time_summary['carcinoma', 'sd_days'], digits = 3)` days).
